### 2017-02-28 paper/chapters/introduction.tex
fixed captions (short caption)
better caption for caption2img
more intro

\begin{figure}[h]
  \includegraphics[width=\linewidth]{media/caption2img}
  \caption[Caption to Image]{Images generated given a captioning text \cite{gan_t2i:2016}}
  \label{fig:caption2img}
\end{figure}
While sampling from generative models has by itself interesting applications, such as producing images given a caption as seen in Figure \ref{fig:caption2img}, there are many applications covering various different areas.
A subset of these applications include
image denoising, % TODO: citation
image impainting, % TODO: citation
text-to-speech, % TODO: citation, e.g. WaveNet
machine translation, % (neural?) TODO: e.g. NMT w/ VAE
and semi-supervised learning, for example discussed in \ref{ssub:cvae}.
\\\\

@@

Generative models are in particular interesting as they

\paragraph{Short-term applications}
\begin{itemize}
  \item \textbf{Data augmentation} is a technique to enrich an existing dataset by adding more derived information to provide more in-depth insight.
  \item \textbf{Semi-supervised learning} on datasets with a small amount of unlabelled data and a large amount of labeled data to improve model performance. Examples include the conditional variational autoencoder discussed in \ref{ssub:cvae}.
  \item \textbf{Data exploration and understanding} is often needed especially for high-dimensional and complex data. By using models which reduce the dimensions by selecting only important features for generative modeling can help to improve data understanding. Figure \ref{fig:vae_experiments} explores the two-dimensional manifold of a handwritten digit, which originally had 784 dimensions.
\end{itemize}

@@

The assumption that the observed variables $x$ can be explained using the causal factors $z$ can be expressed as graphical model shown in Figure \ref{fig:dgm}.
%as having the conditional probability $p(x|z)$ where the latent variables $z$ are the causal factors. This stochastic dependency can be represented in a graphical model shown in figure \ref{fig:dgm}.\\

\begin{figure}[htb]
\centering
\includegraphics{media/directed_graphical_model}
  \caption{Directed graphical model}
  \label{fig:dgm}
  \medskip
  \small
  The latent variable $z$ is a stochastic dependency of the observed data $x$.
\end{figure}

>>>>

%%%
%%% intuition // generative vs discriminative
%%%
Intuitively, generative modeling is based on the idea that in order to produce data similar to previously seen data it helps to understand the data first~\cite{salakhutdinov:2009}.
The case of deep learning suggests the use of deep, hierarchical architectures to capture dependencies within the data and reduce the high-dimensional input~\cite{reducing_dim:2006}.

In contrast to discriminative models which handle the conditional distribution $p(y|x)$ of the labels $y$ given observations $x$, generative models use the joint distribution $p(x,y)$ over both labels and observations to describe the data~\cite{learning_dbn:2006}.\\\\

%TODO more:
% motivation

%%%
%%% applications
%%%

\begin{figure}[h]
  \includegraphics[width=\linewidth]{media/caption2img}
  \caption[Caption-to-Image Samples]{Image Synthesis from provided caption~\cite{gan_t2i:2016}}
\label{fig:caption2img}
\end{figure}
%Generative modeling is a wide field involving capturing an unknown data distribution.\\

% TODO (kinda similar to the first sentences in the introduction (and even abstract as well)
Following the idea of deep models to learn hierarchical representations, these results should be applicable to generative models as well.
% "learning deep generative models" gives these arguments!

Sampling from generative models generally means sampling from a distribution which the model either implicitly or explicitly has learned.
Interesting applications around sampling from these models include generating images given a caption as seen in Figure~\ref{fig:caption2img} or sampling 3D models from a learned latent representation~\cite{gan_3d:2016}.
As generative models often have internal representations of the data distribution, these learned representational structures or features can be used in models built on top.
A subset of these applications include
image denoising~\cite{generative_denoising:2013}, % TODO: citation
text-to-speech, % TODO: citation, e.g. WaveNet
machine translation, % (neural?) TODO: e.g. NMT w/ VAE
and semi-supervised learning, for example discussed in~\ref{ssub:cvae}.
\\\\

\newpage

%%%
%%% history // other models
%%%
\todo[inline]{History}
% traditional models (HMM, GMM)
Traditional models include hidden Markov models used in speech recognition~\cite{hmm:1989} and Gaussian mixture models.

% undirected networks (Boltzmann machine)
Another class in generative models were undirected networks, for example various extensions of the Boltzmann machine~\cite{bm:1985} such as the restricted Boltzmann machine (RBM) and the deep Boltzmann machine~\cite{dbm:2009} were used.\\

% directed networks (SBN, DBN)
For directed networks, the sigmoid belief network (SBN) is an early model which is presented in~\ref{sec:sbn}.
SBN has also been extended using an restricted Boltzmann machine resulting in Deep Belief Nets (DBN)~\cite{learning_dbn:2006}.

@@

\begin{itemize}
  \item \textbf{Data augmentation} is a technique to enrich an existing dataset by adding more derived information to provide more in-depth insight.
  \item \textbf{Semi-supervised learning} on datasets with a small amount of unlabelled data and a large amount of labeled data to improve model performance. Examples include the conditional variational autoencoder discussed in~\ref{ssub:cvae}.
  \item \textbf{Data exploration and understanding} is often needed especially for high-dimensional and complex data. By using models which reduce the dimensions by selecting only important features for generative modeling can help to improve data understanding. Figure~\ref{fig:vae_experiments} explores the two-dimensional manifold of a handwritten digit, which originally had 784 dimensions.
\end{itemize}

@@

The assumption that the observed variables $x$ can be explained using the causal factors $z$ can be expressed as graphical model shown in Figure~\ref{fig:dgm}.
%as having the conditional probability $p(x|z)$ where the latent variables $z$ are the causal factors. This stochastic dependency can be represented in a graphical model shown in figure \ref{fig:dgm}.\\

\begin{figure}[htb]
\centering
\includegraphics{media/directed_graphical_model}
  \caption[Graphical Model]{Directed Graphical Model}
\label{fig:dgm}
  \medskip
  \small
  \todo[inline]{arrow indicate stochastic dependency}
  The latent variable $z$ is a stochastic dependency of the observed data $x$.
\end{figure}
